EDB 2.0 How EAGLE improved the Epigraphic Database Bari

### 2016-02-10 source/28/EAGLE2016_Rocco_engrev.tex
pdfs new folder, fixes to Rocco
removed main.pdf, file too large
updated committee
updated main
updated biblio

%%%%%%%%
%TITLE:%
%%%%%%%%
\title{EDB 2.0 
How EAGLE improved the Epigraphic Database Bari}
\author[uniba]{Anita Rocco \corref{first}}
\address[uniba]{Università degli Studi di Bari ``Aldo Moro''}
\cortext[first]{Corresponding author. Email: anita.rocco@uniba.it}
\begin{document}
\maketitle
\begin{abstract}
This paper is dedicated to the evolution of the Epigraphic Database Bari (EDB) from the minimalistic design of its
origins to its current status. Although EDB, the database of inscriptions by Christians from Rome, dates back to the
late 1980s , involvement in the EAGLE – Europeana project has had a significantly positive impact on its development.
In fact, maintaining its peculiar character, dictated by its own history and, mostly, by the characteristics of its
documentary base, it has taken advantage of the solutions adopted to integrate different archives and purpose-built
best practices. 
\end{abstract}
\keywords{Epigraphic database, EDB, EAGLE Europeana, Christian inscriptions, Late antique inscriptions }

@@

\begin{figure}[hbp]
\centering
 \includegraphics[width=\columnwidth]{Fig1NEW.png} 
\caption{The record of a bilingual inscription (ICVR V, 15147) as it appeared in data processing program ICVR in MS-DOS
in 1988.}
\label{fig:1}
\end{figure}

@@

In the early 2000s, the ICVR database, containing more than 20,000 records, became part of the EAGLE federation of
databases (Electronic Archive of Greek and Latin Epigraphy), under the patronage of the Association Internationale
d’Épigraphie Grecque et Latine (AIEGL), as EDB (Epigraphic Database Bari) and extended its competencies to the
epigraphic documentation of Christian patronage of the city of Rome, published after the volumes of ICVR.

@@

The current structure of EDB consists of a relational database, based on the open-source program My-SQL, with a complex
scheme drafted according to the most recent advances in epigraphic methodology: reestablishing historical and material
value of the object, identifying each epigraph as a complex and polysemic product consisting partially but not only of
text.

@@

Besides data related to the ICVR publication - volume and edition number of the inscription - now EDB is able to record
other bibliographic references as well as concordances with other Corpora (CIL, IG, IGUR, IGC).

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{Fig4.png}
\caption{Bibliographic data fields}
\label{fig:3b}
\end{figure}

@@

Data on the \emph{original context} are therefore organized hierarchically in three related fields containing controlled lists.
After selecting the area of the suburb identified by the name of the consular road - or by the number of the Augustan
\emph{regio} for urban inscriptions - it's possible to select the monument from a list: a catacomb or part of it, if it is a
large and multi-layered one; a church; a public building or an urban area. The third field allows access to a further
associated list where the position of the epigraph inside the monument can be selected. For the catacombs in
particular, it’s possible to use these fields to annotate the gallery or the cubicle, named with the alphanumeric code
used in the maps published in ICVR.

@@

 \begin{figure}[hbp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016Roccoengrev-img006.png} 
\caption{The original context maps.}
\label{fig:6}
\end{figure}

@@

Even the geographic data relating to the place of conservation of the inscriptions are managed with a similar structure.
Since not a few inscriptions, produced in Rome have been taken away and carried to other places in Italy or abroad, the
information have been organized in three related fields reporting respectively the list of the cities, the list of
associated structures, such as museums, churches or catacombs; and the specific positions in the context where the
object is actually preserved.

@@

Adopting the best practice suggested by the EAGLE consortium, georeferencing is guaranteed even for the Conservation
data, linking every City/Town to GeoNames site\footnote{ \url{www.geonames.org}}, which allows the user to pinpoint the
location and to avoid ambiguity between homonyms. A link to the Trismegistos Collection\footnote{
\url{www.trismegistos.org}}, a database of papyrological and epigraphic collections, if available, also helps to identify
uniquely the place of conservation, as well as to obtain additional information, included the geographic
positioning.\footnote{\url{http://www.eagle-network.eu/wp-content/uploads/2013/06/EAGLE\_D2.2.2\_Content-harmonisation-guidelines-including-GIS-and-terminologies-Second-Release.pdf}}

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{Fig8.png}
\caption{Conservation maps and links}
\label{fig:8new}
\end{figure}

@@

\begin{figure}[hbp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016Roccoengrev-img008.png}
\caption{Controlled vocabularies for Type of support, Executing technique and Function fields.}
\label{fig:8}
\end{figure}

The controlled lists have been integrated in the vocabularies of the EAGLE community\footnote{
\url{http://www.eagle-network.eu/resources/vocabularies/}.}, wich align, harmonize, create relations and translate into
various languages the terms used by the various partners, and returns them in a format that allows the user to get a
stable and unique identifier for each term, accessible and reusable by other users.

\begin{figure}[hbp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016Roccoengrev-img009.png}
\caption{The integration with EAGLE Vocabularies, that can be opened clicking on the highlighted terms.}
\label{fig:9}
\end{figure}

@@

\begin{figure}[hbp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016Roccoengrev-img010.png}
\caption{Text fields in the submission mask.}
\label{fig:10}
\end{figure}

@@

In EDB, as in every epigraphic database, the transcription of the text provides a real pre-edition, with systematic
expansion of abbreviations, hypothesis of integration, and, if possible, an update version of the text, according to
recent publications. It includes, moreover, the description of non-alphabetic signs - in double round brackets - such
as figurative elements (fishes, birds, anchors, etc.) and Christological monograms. Other fields record onomastic
notes, \emph{critical apparatus} and textual comments.

@@

@@

Another meaningful improvement of EDB 2.0 is represented by the inclusion of visual representation of inscriptions. It’s
evident how images dramatically enhance analysis of epigraphic materials, showing them in their manifold aspects:
reference with the context, form and quality of the support, graphic forms, peculiarities of technique, layout and
relationship between the text and any figurative or decorative elements\footnote{\citet{panciera_cronistoria_2006}.}.

@@

The latter of these can navigate in the descriptive section of the website (About EDB, People) and in the list of cited
Publications. They also have access to the entire database using two research masks: a \emph{Quick search} allows the user to
search in only one of the following fields: identifier EDB, bibliographic data and text; an \emph{Advanced Search} provides
the opportunity to explore the database through multiple search criteria variously combined.

@@

\begin{enumerate}
\item The inclusion of EDB bibliography, structured in metadata, in the Zotero Group, where it have been merged with those
of other content providers and has been made directly and publicly available in the most reusable way, giving more
exposition of the bibliographic database; allowing the integration and enrichment with other databases; allowing easy
export of data in multiple formats (bibtex, bookmarks, mods, rdf, xml, etc).

>>>>

%%%%%%%%
%TITLE:%
%%%%%%%%
\title{EDB 2.0 
How EAGLE improved the Epigraphic Database Bari}
\author[uniba]{Anita Rocco \corref{first}}
\address[uniba]{Dip. di Studi Umanistici, Università degli Studi di Bari ``Aldo Moro''}
\cortext[first]{Corresponding author. Email: anita.rocco@uniba.it}
\begin{document}
\maketitle
\begin{abstract}
This paper is dedicated to the evolution of the Epigraphic Database Bari (EDB) from the minimalistic design of its
origins to its current status. Although EDB, the database of inscriptions by Christians from Rome, dates back to the
late 1980s , involvement in the EAGLE – Europeana project has had a significantly positive impact on its development.
In fact, maintaining its peculiar character, dictated by its own history and, mostly, by the characteristics of its
documentary base, it has taken advantage of the solutions adopted to integrate different archives and purpose-built
best practices. 
\end{abstract}
\keywords{Epigraphic database, EDB, EAGLE Europeana, Christian inscriptions, Late antique inscriptions }

@@

\begin{figure}[hbp]
\centering
 \includegraphics[width=\columnwidth]{Fig1NEW.png} 
\caption{The record of a bilingual inscription (ICVR V, 15147) as it appeared in data processing program ICVR in MS-DOS
in 1988 and as it appears now in EDB, with the acronyms expanded, the image and the text displayed in Greek characters.}
\label{fig:1}
\end{figure}

@@

In the early 2000s, the ICVR database, containing more than 20,000 records, became part of the EAGLE federation of
databases (Electronic Archive of Greek and Latin Epigraphy), under the patronage of the Association Internationale
d’Épigraphie Grecque et Latine (AIEGL), as EDB (Epigraphic Database Bari) and extended its competencies to the
epigraphic documentation of Christian patronage of the city of Rome, published after the volumes of ICVR (Fig.~\ref{fig:3}).

@@

The current structure of EDB consists of a relational database, based on the open-source program My-SQL, with a complex
scheme drafted according to the most recent advances in epigraphic methodology: reestablishing historical and material
value of the object, identifying each epigraph as a complex and polysemic product consisting partially but not only of
text (Fig.~\ref{fig:4}).

@@

Besides data related to the ICVR publication - volume and edition number of the inscription - now EDB is able to record
other bibliographic references as well as concordances with other Corpora (CIL, IG, IGUR, IGC) (Fig.~\ref{fig:3b}).

\newpage
\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{Fig4.png}
\caption{Bibliographic data fields.}
\label{fig:3b}
\end{figure}

@@

Data on the \emph{original context} are therefore organized hierarchically in three related fields containing controlled lists (Fig.~\ref{fig:5}). 
After selecting the area of the suburb identified by the name of the consular road - or by the number of the Augustan
\emph{regio} for urban inscriptions - it's possible to select the monument from a list: a catacomb or part of it, if it is a
large and multi-layered one; a church; a public building or an urban area. The third field allows access to a further
associated list where the position of the epigraph inside the monument can be selected. For the catacombs in
particular, it’s possible to use these fields to annotate the gallery or the cubicle, named with the alphanumeric code
used in the maps published in ICVR.

@@

@@

 \begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016Roccoengrev-img006.png} 
\caption{The original context maps.}
\label{fig:6}
\end{figure}

Even the geographic data relating to the place of conservation of the inscriptions are managed with a similar structure.
Since not a few inscriptions, produced in Rome have been taken away and carried to other places in Italy or abroad, the
information have been organized in three related fields reporting respectively the list of the cities, the list of
associated structures, such as museums, churches or catacombs; and the specific positions in the context where the
object is actually preserved (Fig.~\ref{fig:7}).

@@

Adopting the best practice suggested by the EAGLE consortium, georeferencing is guaranteed even for the Conservation
data, linking every City/Town to GeoNames site\footnote{ \url{www.geonames.org}}, which allows the user to pinpoint the
location and to avoid ambiguity between homonyms. A link to the Trismegistos Collection\footnote{
\url{www.trismegistos.org}}, a database of papyrological and epigraphic collections, if available, also helps to identify
uniquely the place of conservation, as well as to obtain additional information, included the geographic
positioning (Fig.~\ref{fig:8new}).\footnote{\url{http://www.eagle-network.eu/wp-content/uploads/2013/06/EAGLE\_D2.2.2\_Content-harmonisation-guidelines-including-GIS-and-terminologies-Second-Release.pdf}}

\newpage
\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{Fig8.png}
\caption{Conservation maps and links.}
\label{fig:8new}
\end{figure}

@@

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016Roccoengrev-img008.png}
\caption{Controlled vocabularies for Type of support, Executing technique and Function fields.}
\label{fig:8}
\end{figure}

The controlled lists have been integrated in the vocabularies of the EAGLE community\footnote{
\url{http://www.eagle-network.eu/resources/vocabularies/}.}, wich align, harmonize, create relations and translate into
various languages the terms used by the various partners, and returns them in a format that allows the user to get a
stable and unique identifier for each term, accessible and reusable by other users (Fig.~\ref{fig:9}).

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016Roccoengrev-img009.png}
\caption{The integration with EAGLE Vocabularies, that can be opened clicking on the highlighted terms.}
\label{fig:9}
\end{figure}

\newpage

@@

@@

In EDB, as in every epigraphic database, the transcription of the text provides a real pre-edition, with systematic
expansion of abbreviations, hypothesis of integration, and, if possible, an update version of the text, according to
recent publications. It includes, moreover, the description of non-alphabetic signs - in double round brackets - such
as figurative elements (fishes, birds, anchors, etc.) and Christological monograms. Other fields record onomastic
notes, \emph{critical apparatus} and textual comments (Fig.~\ref{fig:10}).

@@

\newpage

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016Roccoengrev-img010.png}
\caption{Text fields in the submission mask.}
\label{fig:10}
\end{figure}

@@

Another meaningful improvement of EDB 2.0 is represented by the inclusion of visual representation of inscriptions. It’s
evident how images dramatically enhance analysis of epigraphic materials, showing them in their manifold aspects:
reference with the context, form and quality of the support, graphic forms, peculiarities of technique, layout and
relationship between the text and any figurative or decorative elements\footnote{\citet{panciera_questioni_2006}.}.

@@

The latter of these can navigate in the descriptive section of the website (About EDB, People) and in the list of cited
Publications. They also have access to the entire database using two research masks: a \emph{Quick search} allows the user to
search in only one of the following fields: identifier EDB, bibliographic data and text; an \emph{Advanced Search} provides
the opportunity to explore the database through multiple search criteria variously combined (Fig.~\ref{fig:11}).

@@

\begin{enumerate}
\item The inclusion of EDB bibliography, structured in metadata, in the Zotero Group, where it has been merged with those
of other content providers and has been made directly and publicly available in the most reusable way, giving more
exposition of the bibliographic database; allowing the integration and enrichment with other databases; allowing easy
export of data in multiple formats (bibtex, bookmarks, mods, rdf, xml, etc).
